% ============================================================
% Data Memo: Statistik Industri (SI) Panel
% Green-Trade Project — Energy Data Cleaning
% ============================================================
\documentclass[11pt, a4paper]{article}

\usepackage[margin=2.5cm]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{longtable}
\usepackage{hyperref}
\usepackage{parskip}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{microtype}

% Colours
\definecolor{noteblue}{RGB}{30, 90, 160}
\definecolor{warnred}{RGB}{180, 30, 30}
\definecolor{lightgray}{RGB}{245, 245, 245}

% Section formatting
\titleformat{\section}{\large\bfseries\color{noteblue}}{}{0em}{}[\titlerule]
\titleformat{\subsection}{\normalsize\bfseries}{}{0em}{}

% Hyperlink style
\hypersetup{colorlinks=true, linkcolor=noteblue, urlcolor=noteblue}

% ============================================================
\begin{document}

\begin{center}
  {\LARGE \textbf{Data Memo: Statistik Industri Panel}}\\[0.5em]
  {\large Green-Trade Project — Energy Variables}\\[0.3em]
  {\small Last updated: March 2026 \quad|\quad File: \texttt{data8514\_SI.dta}}
\end{center}

\vspace{1em}
\hrule
\vspace{1em}

% ============================================================
\section{Dataset Overview}

The Statistik Industri (SI) — also known as the Industri Besar dan Sedang (IBS) survey — is
Indonesia's annual census of medium and large manufacturing establishments. It is conducted by
BPS (Badan Pusat Statistik) and covers all establishments with 20 or more workers.

\begin{table}[h]
\centering
\begin{tabular}{ll}
\toprule
\textbf{Item} & \textbf{Value} \\
\midrule
File            & \texttt{data8514\_SI.dta} \\
Years covered   & 1985–2014 \\
Total obs.      & 579,533 firm-year observations \\
Unique firms    & 64,950 establishments \\
Provinces       & 38 \\
Variables       & 98 \\
\bottomrule
\end{tabular}
\caption{Dataset summary}
\end{table}

\textbf{Unit of observation:} one row = one establishment in one year.

\textbf{Key identifiers:}
\begin{itemize}[noitemsep]
  \item \texttt{psid} — Plant/establishment ID (panel identifier)
  \item \texttt{year} — Survey year
  \item \texttt{disic5} / \texttt{disic4} — 5-digit and 4-digit industry code (ISIC)
  \item \texttt{dprovi} — Province code (38 provinces)
  \item \texttt{ekspor} — Export status (1 = exporter, 2 = non-exporter; 95,759 non-missing obs)
\end{itemize}


% ============================================================
\section{Energy Variables}

\subsection{Variable Naming Convention}

Energy variables follow a structured naming pattern:

\begin{center}
\texttt{[type][use][fuel][suffix]}
\end{center}

\begin{table}[h]
\centering
\begin{tabular}{lll}
\toprule
\textbf{Position} & \textbf{Code} & \textbf{Meaning} \\
\midrule
\multirow{1}{*}{Type}  & \texttt{z}    & Fuel/energy variable \\
                       & \texttt{ef}   & Electricity (special prefix) \\
\midrule
\multirow{2}{*}{Use}   & \texttt{p}    & For production (\textit{untuk produksi}) \\
                       & \texttt{n}    & Non-production (\textit{bukan untuk produksi}) \\
\midrule
\multirow{6}{*}{Fuel}  & \texttt{d}    & Solar/HSD (diesel) \\
                       & \texttt{s}    & Minyak tanah (kerosene) \\
                       & \texttt{o}    & Bensin/premium (gasoline) \\
                       & \texttt{i}    & Batubara (coal) or industrial fuel \\
                       & \texttt{z}    & Gas/LPG \\
                       & \texttt{x}    & Other fuel (type unclear) \\
\midrule
\multirow{3}{*}{Suffix} & \texttt{vcu} & Value in current Rupiah (Rp000) \\
                        & \texttt{ccu} & Cost in current Rupiah (Rp000) \\
                        & \texttt{kcu} & Physical quantity (units vary by fuel) \\
\bottomrule
\end{tabular}
\caption{Variable naming convention. Fuel type codes are inferred from IBS conventions —
  variable labels are missing from the \texttt{.dta} file. Verify against questionnaire PDFs.}
\end{table}

\noindent\textbf{Note:} Variable labels are \textit{all missing} in the raw \texttt{.dta} file. The fuel type
mappings above are based on standard IBS/SI documentation and need to be confirmed against
the questionnaire PDFs included in the data folder.


\subsection{Electricity}

\begin{table}[h]
\centering
\begin{tabular}{lllll}
\toprule
\textbf{Variable} & \textbf{Description} & \textbf{Coverage} & \textbf{Years} & \textbf{Units} \\
\midrule
\texttt{efuvcu} & Electricity expenditure & 93.6\% (542,583 obs) & 1985–2014 & Rp000 \\
\bottomrule
\end{tabular}
\caption{Electricity variable. Nearly complete across all years. \textit{Not} the sum of
  the fuel variables below — confirmed by low correlation (0.40) and consistently smaller values.}
\end{table}


\subsection{Fuel Variables}

All fuel variables come in pairs: \texttt{zp*} (production use) and \texttt{zn*} (non-production use).
Monetary values (\texttt{vcu}) are the primary series; physical quantities (\texttt{kcu}) are available
only for a subset of observations (see Section~\ref{sec:sparse}).

\begin{longtable}{lllll}
\toprule
\textbf{Variable} & \textbf{Fuel Type} & \textbf{Use} & \textbf{Non-zero obs} & \textbf{Years} \\
\midrule
\endhead
\texttt{zpdvcu} & Diesel/Solar (HSD) & Production     & 515,623 (89.0\%) & 1990–2014 \\
\texttt{zndvcu} & Diesel/Solar (HSD) & Non-production & 417,400 (72.0\%) & 1990–2014 \\
\texttt{zpsvcu} & Kerosene           & Production     & 515,053 (88.9\%) & 1990–2014 \\
\texttt{znsvcu} & Kerosene           & Non-production & 416,404 (71.9\%) & 1990–2014 \\
\texttt{zpovcu} & Gasoline           & Production     & 160,715 (27.7\%) & 1990–2014 \\
\texttt{znovcu} & Gasoline           & Non-production & 109,662 (18.9\%) & 1990–2014 \\
\texttt{zpivcu} & Coal / Ind. fuel   & Production     & 212,132 (36.6\%) & 1990–2014 \\
\texttt{znivcu} & Coal / Ind. fuel   & Non-production & 156,310 (27.0\%) & 1990–2014 \\
\texttt{zpzvcu} & Gas / LPG          & Production     & 172,088 (29.7\%) & 2001–2009 \\
\texttt{znzvcu} & Gas / LPG          & Non-production & 142,421 (24.6\%) & 2001–2009 \\
\texttt{zpxvcu} & Other fuel         & Production     & 43,630  ( 7.5\%) & 2005–2009 \\
\texttt{znxvcu} & Other fuel         & Non-production & 34,302  ( 5.9\%) & 2005–2009 \\
\bottomrule
\caption{Fuel variables — monetary values (\texttt{vcu}). Coverage = share of total 579,533 obs
  with a non-zero, non-missing value.}
\end{longtable}


% ============================================================
\section{Physical Quantity Variables and the Sparsity Problem}
\label{sec:sparse}

Each fuel variable has a physical quantity counterpart (\texttt{kcu} suffix) alongside the
monetary value. These are the variables needed to compute carbon emissions directly. However,
they are very sparse:

\begin{table}[h]
\centering
\begin{tabular}{lrrl}
\toprule
\textbf{Variable} & \textbf{Non-zero obs} & \textbf{\% of 579,533} & \textbf{Likely unit} \\
\midrule
\texttt{zpdkcu} & 66,728 & 11.5\% & Litres (diesel) \\
\texttt{zpskcu} & 34,151 &  5.9\% & Litres (kerosene) \\
\texttt{zpokcu} &  8,877 &  1.5\% & Litres (gasoline) \\
\texttt{zpikcu} & 33,599 &  5.8\% & Tonnes or kg (coal) \\
\texttt{zndkcu} & 40,085 &  6.9\% & Litres (diesel) \\
\texttt{znskcu} & 20,826 &  3.6\% & Litres (kerosene) \\
\texttt{znokcu} &  4,568 &  0.8\% & Litres (gasoline) \\
\texttt{znikcu} & 18,000 &  3.1\% & Tonnes or kg (coal) \\
\bottomrule
\end{tabular}
\caption{Physical quantity (\texttt{kcu}) variables. Available mainly post-2000 for a
  subset of firms. Units are inferred from typical SI survey documentation — confirm
  against questionnaire PDFs.}
\end{table}

\noindent\textbf{Implication for carbon conversion:} Because \texttt{kcu} variables are sparse,
the full-panel carbon emission measure will need to rely on the monetary \texttt{vcu} values
combined with \textbf{fuel price deflators} (BPS/Pertamina/PLN annual price series by fuel
type) to back out physical quantities. This adds a step to the pipeline.


% ============================================================
\section{Planned Cleaning Pipeline}

\begin{enumerate}
  \item \textbf{Verify variable definitions} against questionnaire PDFs in the data folder
        (years 2000–2014 are available).
  \item \textbf{Check cross-year consistency}: the SI questionnaire form changed over time;
        some fuel categories may be merged or split across survey waves.
  \item \textbf{Harmonise monetary values}: confirm units are Rp000; deflate to a base year
        using CPI (variable \texttt{cpi} is present in the dataset).
  \item \textbf{Construct physical quantity series}: use fuel price deflators to convert
        \texttt{vcu} → physical units where \texttt{kcu} is missing.
  \item \textbf{Apply emission conversion factors}: multiply physical quantities by
        fuel-specific $\text{CO}_2$ equivalent factors (source to be documented).
  \item \textbf{Compute carbon intensity}: total $\text{CO}_2$-equivalent / output value.
  \item \textbf{Save clean dataset}: \texttt{output/energy\_clean\_8514.dta}.
\end{enumerate}

\begin{table}[h]
\centering
\begin{tabular}{lll}
\toprule
\textbf{Do-file} & \textbf{Input} & \textbf{Output} \\
\midrule
\texttt{01\_energy\_clean.do}   & \texttt{data8514\_SI.dta} & \texttt{energy\_harmonised.dta} \\
\texttt{02\_carbon\_convert.do} & \texttt{energy\_harmonised.dta} & \texttt{energy\_clean\_8514.dta} \\
\bottomrule
\end{tabular}
\caption{Planned do-file pipeline}
\end{table}


% ============================================================
\section{Open Questions}

\begin{enumerate}
  \item \textbf{Fuel type codes}: Are \texttt{d/s/o/i/z/x} correctly mapped to
        diesel/kerosene/gasoline/coal/gas/other? Must verify against questionnaires.
  \item \textbf{Production vs.\ non-production}: Should carbon intensity use production-use
        energy only (\texttt{zp*}), or total (\texttt{zp*} + \texttt{zn*})?
  \item \textbf{Gas/LPG gap}: \texttt{zpz/znz} only runs 2001–2009. How should
        pre-2001 and post-2009 periods be handled?
  \item \textbf{Electricity units}: Is \texttt{efuvcu} expenditure in Rp000? Does it include
        both self-generated and PLN-purchased electricity?
  \item \textbf{\texttt{ccu} vs.\ \texttt{vcu}}: Are the cost (\texttt{ccu}) and value
        (\texttt{vcu}) variants the same variable recorded in different survey waves, or do
        they measure different things?
\end{enumerate}


% ============================================================
\section*{Document History}
\begin{itemize}[noitemsep]
  \item 2026-03-11 — Initial version, Session 1 data exploration
\end{itemize}

\end{document}
